Hydra Battles and Cie

### 2021-03-02 doc/part-hydras.tex
new exercise
new exercise; minor corrections
small corrections

Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
has any number of sons. Note that, contrary to the computer science tradition, we will show the hydras 
with the heads up and the foot (i.e., the root of the tree) down.
Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}). 
\emph{For a less technical description of hydras, please see 
\url{https://www.smbc-comics.com/comic/hydra}.}

@@

At the first round, Hercules choses to chop off the rightmost head of \texttt{Hy}.
Since this head is near the floor, the hydra loses this head. Let us call 
 \texttt{Hy'} the resulting state of the hydra, represented in Fig.~\vref{fig:Hy-prime}.

Next, assume Hercules choses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one. Fig.~\vref{fig:Hy2} represents the rotten neck in dashed lines, and the part that will be replicated in red. Assume also that the hydra decides to add 4 copies of the red part\footnote{In other words, the replication factor at this round is equal to $4$.}. We obtain a new state \texttt{Hy''} depicted in Fig.~\ref{fig:Hy3}.

@@

\begin{remark}
The upper bound of the considered interval has been chosen to be  $l-1$ and not $l$, in order to simplify some statements and proofs in composition lemmas associated to ordinals of the form $\alpha\times i$ and 
$\omega^\alpha\times i + \beta$.
In order to consider any ordinal below $\epsilon_0$, we consider a special case for $\alpha=0$.
\end{remark}

@@

@@

\begin{definition}
 Let $f$ zand $g$ be two arithmetic  functions; $f$ is said to \emph{dominate} $g$ if $f(p)>g(p)$ for any all sufficiently large $p$.
\end{definition}

@@

\begin{Coqsrc}
Lemma LF3 : dominates_from  2 (F_ 3) (fun  n => iterate exp2 n n).
\end{Coqsrc}
\end{exercise}

@@

\begin{exercise}
Prove the following result~\cite{KS81}(p. 298).

\begin{quote}
   For $n\geq 2$ and $\alpha \geq 3$, $F_\alpha(n+1)\geq 2^{F_\alpha(n)}$.
\end{quote}
\end{exercise}

>>>>

Technically, a \emph{hydra} is just a finite ordered tree, each node of which 
has any number of sons. Note that, contrary to the computer science tradition, we will show the hydras 
with the heads up and the foot (i.e., the root of the tree) down.
Fig.~\ref{fig:Hy} represents such  a hydra, which will be referred to as \texttt{Hy} in our examples (please look at the 
module~\href{../theories/html/hydras.Hydra.Hydra_Examples.html}{Hydra.Hydra\_Examples}). 
\emph{For a less formal description of hydras, please see 
\url{https://www.smbc-comics.com/comic/hydra}.}

@@

At the first round, Hercules choses to chop off the rightmost head of \texttt{Hy}.
Since this head is near the floor, the hydra simply loses this head. Let us call 
 \texttt{Hy'} the resulting state of the hydra, represented in Fig.~\vref{fig:Hy-prime}.

Next, assume Hercules choses to chop off one of the two highest heads of \texttt{Hy'}, for instance the rightmost one. Fig.~\vref{fig:Hy2} represents the broken segment in dashed lines, and the part that will be replicated in red. Assume also that the hydra decides to add 4 copies of the red part\footnote{In other words, the replication factor at this round is equal to $4$.}. We obtain a new state \texttt{Hy''} depicted in Fig.~\ref{fig:Hy3}.

@@

\begin{remark}
The upper bound of the considered interval has been chosen to be  $l-1$ and not $l$, in order to simplify some statements and proofs in composition lemmas associated with the ordinals of the form $\alpha\times i$ and 
$\omega^\alpha\times i + \beta$.
In order to consider any ordinal below $\epsilon_0$, we consider a special case for $\alpha=0$.
\end{remark}

@@

These properties are defined in \coq{} in the library
~\href{../theories/html/hydras.Prelude.Iterates.html}{Prelude.Iterates}.

\index{maths}{Abstract properties of arithmetic functions}
\index{hydras}{Abstract properties of arithmetic functions}

\begin{Coqsrc}
(** ** Abstract properties of arithmetic functions *)

Definition strict_mono f := forall n p,  n < p -> f n < f p.

Definition strict_mono1 f := forall n p,  0 < n < p -> f n < f p.

Definition dominates_from n g f  := forall p, n <= p -> f p < g p.

Definition dominates_strong g f  := {i : nat | dominates_from i g f}.

Definition dominates g f := exists i : nat, dominates_from i g f .

Definition fun_le f g  := forall n:nat,  f n <=  g n.

Infix ">>" := dominates (at level 60).

Infix ">>s" := dominates_strong (at level 60).

Infix "<<=" := fun_le (at level 60).
\end{Coqsrc}

@@

@@

\begin{Coqsrc}
Lemma LF3 : dominates_from  2 (F_ 3) (fun  n => iterate exp2 n n).
\end{Coqsrc}

\emph{You may start this exercise with the file
\url{../exercises/ordinals/F_3.v}.}
\end{exercise}

\begin{exercise}
Prove that, for any $\alpha\geq 3$ and $n\geq 2$,
$F_\alpha(1+n)\geq 2^{F_\alpha(n)}$.

\emph{You may start this exercise with the file
\url{../exercises/ordinals/F_3.v}.}
\end{exercise}

%ici
\index{hydras}{Exercises}

\begin{exercise}
It is tempting to prove a simple property of monotony 
of the function \texttt{F\_}.

\begin{quote}
   Let $\alpha\leq\beta<\epsilon_0$. For any $n\geq 2$,
$F_\alpha(n)\leq F_\beta(n)$. 
\end{quote}
Prove or disprove this statement.

\emph{You may start this exercise with the file
\url{../exercises/ordinals/is_F_monotonous.v}.}
\end{exercise}

@@
